---
# title: "DATA 606 - PROJECT PROPOSAL"
# author: "[Kleber Perez - `r format(Sys.Date(), '%B %d, %Y')`](https://www.linkedin.com/in/kleberperez/)"
output: 
  html_document:
    code_folding: show
    theme: sandstone
    highlight: pygments
    css: ./pcss.css
---

<div class="jumbotron jumbotron-fluid">
  <div class="btn btn-default btn-block">
  <br />
  
  <p><a class="label label-info" href="https://www.linkedin.com/in/kleberperez/" target="_blank" role="button">KLEBER PEREZ</a></p>
  <p class="lead">DATA 606 - PROJECT PROPOSAL <br />
  SOURCE FILES AND RMARKDOWN ON 
   <a class="label label-warning" href="http://rpubs.com/kleberperez/482016" target="_blank" role="button">RPUBS</a>
   AND 
   <a class="label label-success" href="https://github.com/kleberperez1/CUNY-SPS-Data606-Project-Proposal" target="_blank" role="button">GITHUB</a>
  </p>
  
  <br />
  
  </div>
</div>

<br />


### Load Libraries 
```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load libraries

library(kableExtra)
#suppressMessages(suppressWarnings(library(tidyverse)))
library(tidyr)
library(dplyr)
library(psych)
library(stringr)

```


### Data Wrangling 
<div class="alert alert-info" role="alert">

Data source from <a class="label label-default" href="http://www.fivethirtyeight.com" target="_blank" role="button">fivethirtyeight</a> and <a class="label label-default"https://github.com/fivethirtyeight/data/tree/master/college-majors" target="_blank" role="button">github</a>.

</div>

```{r}
# Load data and Subset data by category 
url1 <- "https://raw.githubusercontent.com/kleberperez1/CUNY-SPS-Data606-Project-Proposal/master/all_ages.csv"
all_ages <- url1 %>% 
  read.csv(stringsAsFactors = FALSE) %>% 
  tbl_df() %>% 
  arrange(Major_category)

kable(head(all_ages, 20)) %>%
  kable_styling("striped", "hovered", font_size = 12) %>%
  scroll_box(height = "500px")

```

### Subset by Category
```{r}
# Subsetting by Major Category 
all_ages_ag <- all_ages %>% 
  filter(Major_category == "Agriculture & Natural Resources")

kable(head(all_ages_ag, 20)) %>%
  kable_styling("striped", "hovered", font_size = 12) %>%
  scroll_box(height = "500px")

```

### Apply Filter 
```{r}
# Apply Filter on category 
value_list <- c("Arts", "Biology & Life Science", "Business") #, "Communications & Journalism", "Computers & Mathematics", 
                #"Education", "Engineering", "Health", "Humanities & Liberal Arts", "Industrial Arts & Consumer Services", 
                #"Law & Public Policy", "Physical Sciences", "Psychology & Social Work", "Social Science")
all_ages_value <- all_ages %>% 
  filter(Major_category %in% value_list)

kable(all_ages_value) %>%
  kable_styling("striped", "hovered", font_size = 12) %>%
  scroll_box(height = "500px")

```

### Graduate Students 
```{r}
# Load graduate students file and subset data

url2 <- "https://raw.githubusercontent.com/kleberperez1/CUNY-SPS-Data606-Project-Proposal/master/grad_students.csv"
grad_stdnt <- url2 %>% read.csv(stringsAsFactors = FALSE) %>% tbl_df() %>% arrange(Major_category)

grad_ag <- grad_stdnt %>% 
  filter(Major_category == "Agriculture & Natural Resources")

kable(head(grad_ag, 20)) %>%
  kable_styling("striped", "hovered", font_size = 12) %>%
  scroll_box(height = "500px")

```

### Recent Graduates 
```{r}
# Load graduate students file and subset data
url3 <- "https://raw.githubusercontent.com/kleberperez1/CUNY-SPS-Data606-Project-Proposal/master/recent_grads.csv"
rct_grad <- url3 %>% read.csv(stringsAsFactors = FALSE) %>% tbl_df() %>% arrange(Major_category)

rct_ag <- rct_grad %>% 
  filter(Major_category == "Agriculture & Natural Resources")

kable(head(rct_ag, 20)) %>%
  kable_styling("striped", "hovered", font_size = 12) %>%
  scroll_box(height = "500px")

```

<!-- <div class="jumbotron jumbotron-fluid"> -->
<!--   <div class="btn btn-default btn-block"> -->

### Research question 
<div class="alert alert-info" role="alert">
Which college majors offer the best opportunities in terms of unemployment rate and salary?
</div>

### Cases 
<div class="alert alert-info" role="alert">

All_ages - 
Each case represents majors offered by colleges and universities in the US. There are 173 majors represented. These data include both undergrads and grad students.
<br />

Grad Students - 
Each case represents majors offered by colleges and universities in the US. There are 173 majors represented. These data include only grad students aged 25+ years.
<br />

Recent Grads - 
Each case represents majors offered by colleges and universities in the US. There are 173 majors represented. These data include only undergraduate students aged < 28 years. These data also include gender statistics.

</div>

### Data collection 
<div class="alert alert-info" role="alert">

These Data were collated by the <a class="label label-default" href="http://www.fivethirtyeight.com" target="_blank" role="button">fivethirtyeight</a> website and was posted to their github page <a class="label label-default"  href="https://github.com/fivethirtyeight/data/tree/master/college-major" target="_blank" role="button">college major</a>. 

</div>

### Type of study 
<div class="alert alert-info" role="alert">

This is an observational Study

</div>

### Explanatory Variables
<div class="alert alert-info" role="alert">

The explanatory variables are the counts of employed and unemployed college degree holders and the statistics of their income. These data are numerical.

</div>

### Relevant summary statistics 
<div class="alert alert-info" role="alert">

First we will look at overall unemployment rate for the 3 categories: all ages, recent grads, and grad students.

</div>

```{r}
summary(all_ages$Unemployment_rate)
summary(rct_grad$Unemployment_rate)
summary(grad_stdnt$Grad_unemployment_rate)
DT::datatable(summary(all_ages), options = list(pageLength = 7)) 
DT::datatable(summary(all_ages_ag), list(pageLength = 7)) 
DT::datatable(summary(rct_grad), list(pageLength = 7)) 

```


### Unemployment Rate 

```{r}
unempl <- cbind(all_ages$Unemployment_rate, rct_grad$Unemployment_rate, grad_stdnt$Grad_unemployment_rate)
barplot(unempl/nrow(unempl), names.arg = c("All", "Recent Grad", "Grad Student"), 
        xlab = "Unemployment Rate", col = heat.colors(nrow(unempl)))

```

<div class="alert alert-info" role="alert">

It appears that people holding only a Bachelor's degree have nearly twice as high median unemployment as those with higher degrees. <br />
We will also look at median income for the three categories.

</div>

### Median Income by Major All Ages (USD)

```{r}
summary(all_ages$Median)
colors = c("red", "yellow", "green", "violet", "orange", "blue", "pink", "cyan")
hist(all_ages$Median, main = "Histogram for Median Income All Ages", 
     xlab = "Median Income by Major All Ages (USD)", col = colors)

```

### Median Income by Major Recent Graduates (USD)

```{r}
options(scipen = 999) 
summary(rct_grad$Median)
hist(rct_grad$Median, main = "Histogram for Median Income Recent Grads", 
     xlab = "Median Income by Major Recent Grads (USD)", col = colors)

```

### Median Income by Graduate Students (USD)

```{r}
summary(grad_stdnt$Grad_median)
hist(grad_stdnt$Grad_median, main = "Histogram for Median Income Grad Students", 
     xlab = "Median Income by Major Grad Student (USD)", col = colors)

```

### All, Recent Graduate and Graduate Students Median Salary
```{r}
medsal <- cbind(all_ages$Median, rct_grad$Median, grad_stdnt$Grad_median)
barplot(medsal/nrow(medsal), names.arg = c("All", "Recent Grad", "Grad Student"), 
        xlab = "Median Salary", col = heat.colors(nrow(unempl)))

```


<br /><br /><br /><br />


<div class="jumbotron jumbotron-fluid">
  <div class="btn btn-default btn-block">
  <br />
  
  <p><a class="label label-info" href="https://www.linkedin.com/in/kleberperez/" target="_blank" role="button">KLEBER PEREZ</a></p>
  <p class="lead">DATA 606 - PROJECT PROPOSAL <br />
  SOURCE FILES AND RMARKDOWN ON 
   <a class="label label-warning" href="http://rpubs.com/kleberperez/482016" target="_blank" role="button">RPUBS</a>
   AND 
   <a class="label label-success" href="https://github.com/kleberperez1/CUNY-SPS-Data606-Project-Proposal" target="_blank" role="button">GITHUB</a>
  </p>
  
  <br />
  
  </div>
</div>

<br /><br />

> *Please email to: * <a href="mailto:kleber.perez@live.com">Kleber Perez</a> *for any suggestion.*

<footer class="border-top text-muted fixed-bottom">

        DATA 606 PROJECT PROPOSAL - MSDS 2019 Program.

</footer>

\textbf{Text ignored: END of other links}

<br />
<br />
<br />











